---
title: "This is a replica of all the R scripts"
author: "Jianying"
date: "Last complied `r format(Sys.time(), '%d %B, %Y')`"
output: 
  pdf_document:
    latex_engine: xelatex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(flextable)
library(haven)
library(tidyverse)
```

# Getting sample ADaM data from PHUSE 

```{r ADaM}
adam_path <- "https://github.com/phuse-org/TestDataFactory/raw/main/Updated/TDF_ADaM/"
adsl <- haven::read_xpt(paste0(adam_path, "adsl.xpt"))

# Select a few variables
adsl <- adsl %>%
  dplyr::select(STUDYID, USUBJID, SUBJID, AGE, TRT01P, TRTSDT, TRTEDT, RACE, SEX, DISCONFL)

```

# Take a quick peek

```{r quick peek}
str(adsl)
summary(adsl)

table(adsl$SEX)

adsl %>%
  count(SEX)

#for multiple variables
table(adsl$SEX, adsl$TRT01P, adsl$RACE) 

adsl %>%
  count(SEX, TRT01P, RACE)
```

# R processing that matches SAS 

```{r basic}

# to keep (only) SUBJID
adsl %>%
  select(SUBJID) %>%
  head() # keep only first 6 rows

# to keep all variables except SUBJID
adsl %>%
  select(-SUBJID) %>%
  head()

# select female

adsl %>%
  filter(SEX == "F") %>%
  head() 

# multiple conditions:
adsl %>%
  filter(SEX == "F" & AGE > 70) %>%
  head() 

# ascending
adsl %>%
  arrange(AGE)%>%
  head() 

# descending
adsl %>%
  arrange(-AGE)%>%
  head() 

## with case_when function

adsl %>%
  mutate(AGEGR1 = case_when(
    AGE > 50 ~"> 50 years old",
    AGE <= 50 ~"<= 50 years old",
  )) %>%
  head()


```
# Handling missing values
```{r missingVal}
# Unlike SAS, space(s) != blank string in R, 
" " == ""

# This is true 
"" == ""

# is.na to check if it's missing
c(" ", "Y", NA_character_) %>% is.na()

# NA is non-comparable, result in NA
c(" ", "Y", NA_character_) != "Y"

# add one more row to the data with missing AGE and DISCONFL, select the first 3 rows
adsl_na_example <- adsl %>% 
  add_row(SUBJID = "1", DISCONFL = NA_character_, .before = T) %>% 
  slice(1:3) %>% 
  select(SUBJID, DISCONFL)
# demo data
adsl_na_example %>% head()

adsl_na_example %>% filter(DISCONFL != "Y")

# include is.na(DISCONFL) in the filter
adsl_na_example %>% filter(DISCONFL != "Y" | is.na(DISCONFL))

## continue with sortingh

# add one more row to the data with missing AGE and DISCONFL 
# select the first 3 rows
adsl_na_sort <- adsl %>% 
  add_row(SUBJID = "1", AGE = NA_integer_, .before = T) %>% 
  slice(1:5) %>% 
  select(SUBJID, AGE)
  
# ascending
adsl_na_sort %>%
  arrange(AGE)
  
# to be consistent with SAS - NA at the top when ascending
adsl_na_sort %>%
  arrange(!is.na(AGE), AGE)
  
SUBJID <- sample(adsl$SUBJID, 100, replace = FALSE)
WEIGHT <- sample(50:100, 100, replace = TRUE)

adsl_weight <- data.frame(SUBJID, WEIGHT)

# inner join
adsl%>%
  inner_join(adsl_weight, by = "SUBJID") 

# outer join
adsl %>%
  full_join(adsl_weight, by = "SUBJID")

# left join
adsl %>%
  left_join(adsl_weight, by = "SUBJID")
  
# first 5 records of adsl
adsl_1 <- adsl[c(1:5),]

# the next 5 records of adsl
adsl_2 <- adsl[c(6:10),]

# tidyverse method - bind_rows()
adsl_stacked <- adsl_1 %>%
  bind_rows(adsl_2) 

```

# Clinical trial data

```{r dataSource}
library(help = "datasets")
data("ToothGrowth")

library(haven)

adsl <- read_xpt("https://github.com/phuse-org/TestDataFactory/raw/main/Updated/TDF_ADaM/adsl.xpt")
adsl %>%  dim()

```


# Work on ADaM data and build table with Tplyr

```{r Tplyr_flextable }



adsl <- read_xpt("https://github.com/phuse-org/TestDataFactory/raw/main/Updated/TDF_ADaM/adsl.xpt")

## string formatting with Tplyr

library(Tplyr)

options(
  
  # Categorical variable defaults
  tplyr.count_layer_default_formats =
    list(n_counts = f_str("xxx [xx.xx%]", n, pct)),
  
  # Continuous variable defaults
  tplyr.desc_layer_default_formats =
    list('N'         = f_str('xx', n),
         'Mean [SD]' = f_str('xx.xx [xx.xxx]', mean, sd),
         'Median'    = f_str('xx.x', median),
         'Min, Max'  = f_str('xx, xx', min, max))
)


library(dplyr)

# Initiate Tplyr, specify treatment variable, optional where condition

my_table <-  tplyr_table(adsl, TRT01P, where = SAFFL == "Y") %>%
  
  # Add a total group column
  add_total_group() %>%
  
  # Add individual variables here
  add_layer(group_desc(AGE, b = "Age (years)")) %>%
  add_layer(group_count(AGEGR1, b = "Age Group 1 (years)")) %>%
  add_layer(group_count(SEX, b = "Gender")) %>%
  add_layer(group_count(ETHNIC, b = "Ethnicity")) %>%
  add_layer(group_desc(BMIBL, b = "Baseline Body Mass Index (kg/m2)")) %>%
  
  # Build
  build()

head(my_table, n = 9)


my_table <- my_table %>%
  
  # remove repeating labels
  apply_row_masks(., row_breaks = TRUE) %>%
  
  # specify order of relevant variables
  select(row_label1, 
         row_label2,
         `var1_Xanomeline High Dose`,
         `var1_Xanomeline Low Dose`,
         var1_Placebo,
         var1_Total)  

library(flextable)

## a basic flextable
my_flextable <- my_table %>%
  # start flextable 
  flextable() %>%
  autofit()

# my_flextable 

## a nicer flextable
my_flextable <- my_table %>%
  
  # start flextable 
  flextable() %>% 
  autofit() %>%
  
  # add some padding between rows
  padding(padding = 0.5) %>%
  
  # adjust width of first two columns
  width(j = 1:2, width = 0.5) %>%
  
  # align treatment columns to center
  align(part = "all", align = "center", j = 3:6) %>%
  
  # column header labels
  set_header_labels(., values = list(
    row_label1 = 'Variable',
    row_label2 = ' ',
    `var1_Xanomeline High Dose` = 'Xanomeline \nHigh Dose',
    `var1_Xanomeline Low Dose` = 'Xanomeline \nLow Dose',
    var1_Placebo = 'Placebo',
    var1_Total = 'Total')) %>%
  
  # header + footers
  add_header_lines(values = "Table: Demographics (Safety Analysis Set)") %>%
  add_footer_lines(values = "This was produced in R!") %>%
  
  # font size, font name
  fontsize(part = "all", size = 8)
# font()
# font(part = "all", fontname = "Times")  
# This errors, perhaps version issue.  -- SZ

ft <- my_flextable
ft <- autofit(ft)
# ft

ft <- set_table_properties(ft, width = .5, layout = "autofit")
ft

library(officer)

# send the flextable to .docx output
read_docx() %>%
  body_add_flextable(ft) %>%
  print(target = "my_document.docx")


```

# Produce figures

```{r plots}
# Packages
library(haven)
library(ggplot2)

# Data
adam_path <- "https://github.com/phuse-org/TestDataFactory/raw/main/Updated/TDF_ADaM/"
adsl <- as.data.frame(haven::read_xpt(paste0(adam_path, "adsl.xpt")))

# basic scatter plot of weight vs height
my_scatter_plot <- ggplot(adsl, aes(x = HEIGHTBL, y = WEIGHTBL)) +
  geom_point()
my_scatter_plot

# save plot as a png file
ggsave(plot = my_scatter_plot, 
       filename = "my_plot_output_1.png", 
       width = 7, 
       height = 5, 
       units = "in")

png(filename = "my_plot_output_2.png", 
    width = 7, 
    height = 5, 
    units = "in", 
    res = 300)

print(my_scatter_plot)
dev.off()

```

# phuse data for plot

```{r swim plot}
# Packages
library(haven)
library(swimplot)
library(dplyr)

# Data
adam_path <- "https://github.com/phuse-org/TestDataFactory/raw/main/Updated/TDF_ADaM/"
adsl <- haven::read_xpt(paste0(adam_path, "adsl.xpt"))
adae <- haven::read_xpt(paste0(adam_path, "adae.xpt"))
adtte <- haven::read_xpt(paste0(adam_path, "adtte.xpt"))

adsl_new <- adsl %>%
  select(USUBJID, ARM, TRTDURD, SEX) %>%
  slice(1:50) 
adae_new <- adae %>%
  select(USUBJID, AEDECOD, AESEV, AEREL, ASTDY) %>%
  filter(USUBJID %in% adsl_new$USUBJID & ASTDY >= 0)
adtte_new <- adtte %>%
  select(USUBJID, EVNTDESC, AVAL) %>%
  filter(USUBJID %in% adsl_new$USUBJID & EVNTDESC != "Study Completion Date")  
random_duration_of_events <- sample(1:25, nrow(adtte_new), replace = T)
adtte_new <- adtte_new %>% 
  bind_cols(random_duration_of_events) %>%
  mutate(Resp_end = AVAL + random_duration_of_events )
adsl_new <- as.data.frame(adsl_new)
adae_new <- as.data.frame(adae_new)
adtte_new <- as.data.frame(adtte_new)

swimmer_plot(df=adsl_new,
             id='USUBJID',
             end='TRTDURD',
             fill='lightblue',
             width=.85)

arm_plot <- swimmer_plot(df=adsl_new,
                         id='USUBJID',
                         end='TRTDURD',
                         name_fill='ARM',
                         id_order='ARM',
                         col="black",
                         alpha=0.75,
                         width=.8)
arm_plot

swim_plot_stratify <-swimmer_plot(df=adsl_new,
                                  id='USUBJID',
                                  end='TRTDURD',
                                  name_fill='ARM',
                                  col="black",
                                  alpha=0.75,
                                  width=.8,
                                  base_size=14,
                                  stratify= c('SEX'))
swim_plot_stratify

AE_plot <- arm_plot + 
  swimmer_points(df_points=adae_new,
                 id='USUBJID',
                 time='ASTDY',
                 name_shape='AESEV',
                 size=2.5,
                 fill='white', 
                 name_col='AEREL')
AE_plot

Response_plot <- arm_plot +
  swimmer_lines(df_lines=adtte_new,
                id='USUBJID',
                start ='AVAL',
                end='Resp_end',
                name_col='EVNTDESC',
                size=1)
Response_plot

Response_plot_with_points <- Response_plot +
  swimmer_points_from_lines(df_lines=adtte_new,
                            id='USUBJID',
                            start='AVAL',
                            end='Resp_end',
                            name_col='EVNTDESC',
                            size=2) +
  scale_fill_manual(name="Treatment",
                    values=c("Placebo" ="#A9342F", 
                             "Xanomeline High Dose"="#5B7ACE",
                             "Xanomeline Low Dose"='#FFC300'))+
  scale_color_manual(name="Response",
                     values=c("grey20"))+
  scale_shape_manual(name='',
                     values=c(17,15),
                     breaks=c('AVAL','Resp_end'),
                     labels=c('Response start','Response end'))+
  guides(fill = guide_legend(override.aes = list(shape = NA))) +
  scale_y_continuous(name = "Time since enrollment")

Response_plot_with_points

# Packages
library(gridExtra)
library(grid)

# Data
wp <- data.frame(subjidn = 1:30,
                 trtp = sample(c('Drug','Placebo'), replace = T, 30),
                 aval = runif(30, min = -40, max = 40))

wp %>%  head()

basic_waterfall <- ggplot(wp, aes(y = aval,x = reorder(subjidn, -aval))) +
  geom_bar(stat = "identity")

basic_waterfall

custom_waterfall <- ggplot(wp, aes(y = aval, x = reorder(subjidn, -aval), fill = trtp)) +                            
  geom_bar(stat = "identity") +
  scale_fill_manual("Planned Treatment Arm", values = c('#00bbff','#89d329')) +
  scale_y_continuous(limits = c(-40,40), breaks = seq(-40, 40, by = 10)) +      
  ylab("Maximum reduction of tumor size (%)\n") +                              
  theme_light() +                                                               
  theme(axis.title.x = element_blank(),                                         
        axis.line.x  = element_blank(),
        axis.text.x  = element_blank(),
        axis.ticks.x = element_blank(),
        legend.position = "bottom")                                             

custom_waterfall



```


# on a simulated data

```{r simulated data plot}

set.seed(100)  

n <- 120   # size/records of simulated data

dat_all <- 
  data.frame(SUBJID = 120010100:(120010100+n-1),
             AVAL = c(rnorm(round(0.8*n), 0, 20), rep(0, round(0.2*n))) %>% sample(),
             OVERALLRESP = c("PR", "SD","PD") %>% sample(size=n, replace=TRUE),
             AMEDGRPN = seq(10, 50, 10) %>% sample(size=n, replace=TRUE),
             DOR = rpois(n, 40),
             AVISIT = c("C1", "C2", "C3","EOT") %>% sample(size=n, replace=TRUE),
             ATMLOSS_L = c("E", "L", "Data unavailable") %>% sample(size=n, replace=TRUE),
             ATMLOSS_P = c("E", "L", "Data unavailable") %>% sample(size=n, replace=TRUE),
             ATMLOSS_R = c("E", "L", "Data unavailable") %>% sample(size=n, replace=TRUE),
             IDFOOT = "Bayer: /study/path/.../f_waterfall.sas               30NOV2021 16:06"
  ) %>%
  mutate_at(vars("OVERALLRESP", "DOR", "IDFOOT"), as.character)

waterfall.plot <- dat_all %>% ggplot(aes(reorder(SUBJID, -AVAL), AVAL, fill =OVERALLRESP)) + 
  geom_bar(stat="identity") +
  labs(x = "Subject", 
       y = "Best % Change from baseline \n in sum of diameters \n (target lesions)\n") +
  scale_x_discrete(labels = function(x) stringr::str_wrap(gsub("([0-9])([0-9])", "\\1 \\2 ", x), 
                                                          width = 5) %>% 
                     stringr::str_replace_all(" ", "")) +
  scale_fill_manual("Best Response",
                    breaks = c("PR", "SD","PD"),
                    values=c("PR"='#89d329',
                             "SD"="#756bb1",
                             "PD"='#00bbff')) + 
  geom_text(aes(label = if_else(AVAL == 0,OVERALLRESP,""),fontface="bold"), 
            vjust = -1,
            size=3,
            color="black") + theme_bw() +     
  theme(
    axis.text = element_text(size=12,color="black",face = "bold"),  
    axis.title.y = element_text(size=12, face="bold"), 
    axis.title.x = element_blank(),  
    legend.background = element_rect(color = "steelblue", linetype = "solid"),
    legend.justification = c(1, 1), 
    legend.position = c(1, 1),
    legend.direction = "horizontal",
    legend.text = element_text(size=8, color = "black", face = "bold"),
    legend.title = element_text(size=8, color = "black", face = "bold"),  
    plot.caption = element_text(hjust = 0, size = 10, color = "blue"), 
    plot.caption.position =  "plot"
  ) 

waterfall.plot


var <- c("DOR", "AVISIT")
var_label <- c("Duration (days)", "Timepoint")

add.plot1 <- dat_all %>%
  reshape2::melt(measure.vars = eval(var), value.name = "label", variable.name = "layer") %>%
  mutate(layer = factor(layer, levels = var, labels = var_label)) %>%
  ggplot(aes(reorder(SUBJID, -AVAL))) +
  geom_text(aes(y = layer, label = label), size = 3, fontface = "bold") +
  labs(y = "", x = NULL) +
  theme_minimal() +
  theme(
    axis.text.y = element_text(
      size = 10,
      colour = "black",
      face = "bold"
    ),
    axis.line = element_blank(),
    axis.ticks = element_blank(),
    axis.text.x = element_blank(),
    panel.grid = element_blank(),
    strip.text = element_blank()
  ) +
  coord_fixed(ratio = .8)
add.plot1

footnote1 <- 'Footnote1: "\\n\" can be used to split long footnote into next line.'
footnote2 <- "Footnote2: To avoid manually adding split character. The function stringr::str_wrap   can be used here to wrap the long text with a specified width."
footnote <- lapply(c(footnote1,
                     footnote2,
                     "",
                     dat_all$IDFOOT[1]),
                   function(x) stringr::str_wrap(x, width=120)) %>% # apply str_wrap to individual footnote
  unlist() %>%       # convert list structure to vector
  stringr::str_flatten('\n')  # add split character(new line) between footnotes

var <- c("ATMLOSS_L", "ATMLOSS_P","ATMLOSS_R")
var_label <- c("Local", "Prospective", "Retrospective")

add.plot2 <- dat_all %>%  
  reshape2::melt(measure.vars = eval(var), 
                 value.name = "label",
                 variable.name = "layer") %>%
  mutate(label=case_when(
    label == "L" ~ "ATM Loss",
    label == "E" ~ "ATM Expressed",      
    label == "9" ~ "Data unavailable",
    TRUE ~ label
  )) %>% 
  mutate(layer = factor(layer, levels = rev(var), labels = rev(var_label))) %>% 
  ggplot() +
  aes(reorder(SUBJID, -AVAL), layer, color=label,shape=label) +
  geom_point(size=3)+
  scale_shape_manual(breaks = c("ATM Loss","ATM Expressed", "Data unavailable"), 
                     values = c("ATM Loss"=15,"ATM Expressed"=0, 
                                "Data unavailable"=7))+
  scale_color_manual(values = c("ATM Loss"="black", "ATM Expressed"="black", 
                                "Data unavailable"= 'black'))+
  theme_classic()+
  theme(axis.text=element_text(size=10, colour = "black",face = "bold"),
        axis.title=element_blank(),
        axis.line = element_blank(),
        axis.ticks = element_blank(),
        axis.text.x = element_blank(),
        legend.title = element_blank(),
        legend.text = element_text(size=10, color = "black", face = "bold"),
        legend.position = "bottom",
        panel.border = element_blank(),
        panel.grid = element_blank(), 
        strip.text = element_blank(),
        plot.caption = element_text(hjust = 0, size = 10), 
        plot.caption.position =  "plot"
  )+
  coord_fixed(ratio=.9)+ 
  labs(caption = footnote)
add.plot2

p1 <- waterfall.plot %>% ggplotGrob()
p2 <- add.plot1 %>% ggplotGrob()
p3 <- add.plot2 %>% ggplotGrob()
gtable_rbind(p1, p2, p3,
             size='first') %>% grid.draw()


```
# The so called spider plot, really the clinical endpoint over time

```{r spider plot}

# Packages
library(ggrepel)

# Data
sp <- data.frame(subjidn = rep(1:5, times = c(5,4,5,3,4)),
                 trtp = rep(c('drug','placebo'), times = c(8,13)),
                 avisit = c(0:4,0:3, 0:4, 0:2, 0:3),
                 aval = c(0,5,15,20,20,
                          0,-10,-20,-15,
                          0,-30,-25,-35,-30,
                          0,15,30,
                          0,-20,-15,-20))

sp

basic_spider <- ggplot(sp, aes(x = avisit, y = aval, group = subjidn)) +
  geom_line()

basic_spider

custom_spider <- ggplot(sp, aes(x = avisit, y = aval, group = subjidn, color = trtp)) +                             
  geom_line() +
  geom_point() +
  geom_hline(yintercept = 0, linetype = "dashed") +
  scale_color_manual("Planned Treatment Arm", values = c('#00bbff','#89d329')) +
  scale_y_continuous(limits = c(-40,40), breaks = seq(-40, 40, by = 10)) +      
  ylab("Change in Target Lesion from Baseline (%)\n") +                         
  xlab("Time (weeks)") +                                                        
  theme_light() +                                                               
  theme(legend.position = "bottom")                                             
  
custom_spider

sp_ends <- sp %>% 
  group_by(subjidn) %>% 
  top_n(1, avisit) 

custom_spider +
  geom_text_repel(
    aes(label = subjidn),
    color="black",
    data=sp_ends,
    size = 3,
    direction = "x",
    hjust = 1
  )



```
# Now, let's look at the survival analysis and its plot

```{r survival}

# Packages
library(survminer)
library(survival)
library(broom)
library(flextable)

# Data
adtte <- haven::read_xpt(
  paste0("https://github.com/phuse-org/TestDataFactory/",
         "raw/main/Updated/TDF_ADaM/adtte.xpt"))
         

surv_model <- adtte %>%
  filter(PARAMCD == "TTDE") %>%
  select(STUDYID, USUBJID, PARAMCD, AVAL, CNSR, TRTA) %>%
  survfit(Surv(AVAL, 1-CNSR) ~ TRTA, data = .)

head(tidy(surv_model))
ggsurvplot(
   fit = surv_model,
   data = adtte,
   risk.table = TRUE,
   #ylab = ylabs,
   xlab = "Months",
   linetype = "strata",
   conf.int = F,
   legend.title = "Randomized Treatment",
   legend = c(0.1, 0.1),
   #palette = c(color_trt1,color_trt2),
   risk.table.title = "Number of subjects at risk",
   risk.table.y.text = F,
   risk.table.height = .15,
   censor.shape = 1,
   censor.size = 3,
   ncensor.plot = F,
   xlim = c(0,250),
   xscale = "d_m",
   break.x.by = 30.4375,
   break.y.by = .1,
   ggtheme = theme_survminer(
     font.main = c(10, "bold"),
     font.submain = c(10, "bold"),
     font.x = c(12, "bold"),
     font.y = c(12, "bold"),
   ) + theme(panel.border = element_rect(fill = NA)),
   tables.theme = theme_cleantable()
 )

ggsurvplot(
   fit = surv_model,
   data = adtte,
   risk.table = TRUE,
   #ylab = ylabs,
   xlab = "Months",
   linetype = "strata",
   conf.int = T,
   legend.title = "Randomized Treatment",
   legend = c(0.1, 0.1),
   #palette = c(color_trt1,color_trt2),
   risk.table.title = "Number of subjects at risk",
   risk.table.y.text = F,
   risk.table.height = .15,
   censor.shape = 1,
   censor.size = 3,
   ncensor.plot = F,
   #xlim = c(0,250),
   xscale = "d_m",
   break.x.by = 30.4375,
   break.y.by = .1,
   ggtheme = theme_survminer(
     font.main = c(10, "bold"),
     font.submain = c(10, "bold"),
     font.x = c(12, "bold"),
     font.y = c(12, "bold"),
   ) + theme(panel.border = element_rect(fill = NA)),
   tables.theme = theme_cleantable()
 )

```
# FDA likes the scatter plot for safety

```{r scatter}

# Packages
library(ggplot2)
library(dplyr)

# Data
advs <- haven::read_xpt(
  paste0("https://github.com/phuse-org/TestDataFactory/",
         "raw/main/Updated/TDF_ADaM/advs.xpt"))

## extract information from this dataset

# This  step is to obtain baseline values for each USUBJID for the parameter and analysis time point of interest.
baseline <- advs %>%
  filter(
    SAFFL == "Y",
    PARAMCD == "SYSBP",
    ATPT == "AFTER LYING DOWN FOR 5 MINUTES",
    TRTA != "Placebo"
  ) %>%
  dplyr::select(USUBJID, TRTA, BASE) %>%
  distinct(USUBJID, .keep_all = TRUE)

# The next step is to obtain the maximum value for each USUBJID post-baseline for the same parameter and analysis time point of interest.

post <- advs %>%
  filter(
    SAFFL == "Y",
    PARAMCD == "SYSBP",
    ATPT == "AFTER LYING DOWN FOR 5 MINUTES",
    TRTA != "Placebo",
    AVISIT != "Baseline",
    ANL01FL == "Y"
  ) %>%
  dplyr::select(USUBJID, AVAL) %>%
  group_by(USUBJID) %>%
  arrange(desc(AVAL)) %>%
  slice(1) %>%
  ungroup()

## use left_join to create the data
all <- baseline %>%
  left_join(post)

ggplot(data = all, aes(x = BASE, y = AVAL)) +
  geom_point()


yaxis_text <- "\nMaximum Post-baseline Systolic Blood Pressure (mmHg)\n"
xaxis_text <- "\nBaseline Systolic Blood Pressure (mmHg)"
header <- "Figure 17. Baseline vs. Maximum Systolic Blood Pressure by Treatment Arm, Safety Population¹, Pooled Analysis\n"
footer1 <- "Source: [include Applicant source, datasets and/or software tools used]."
footer2 <- "¹Gray dotted line = no increase; blue line = high-dose treatment linear regression; orange dashed line = low-dose linear regression."

f17 <- ggplot(all, aes(x = BASE, y = AVAL, linetype = TRTA)) +
  geom_point(colour = "black", shape = 21, size = 4, alpha = 0.6, aes(fill = TRTA)) +
  stat_smooth(method = "lm", se = FALSE, aes(color = TRTA)) +
  scale_color_manual(values = c("skyblue", "orange")) +
  scale_fill_manual(values = c("skyblue", "orange")) +
  geom_abline(intercept = 0, slope = 1, size = 1, lty = "dotted", color = "#5A5A5A") +
  scale_x_continuous(breaks = seq(90, 210, 30), limits = c(90, 210)) +
  scale_y_continuous(breaks = seq(90, 210, 30), limits = c(90, 210)) +
  labs(
    y = yaxis_text,
    x = xaxis_text,
    title = header,
    caption = paste(footer1, footer2, sep = "\n")
  ) +
  theme_light() +
  theme(
    panel.background = element_rect(fill = NA, color = "skyblue3", size = 2, linetype = "solid"),
    plot.caption = element_text(hjust = 0),
    legend.position = "bottom",
    legend.title = element_blank(),
    axis.text = element_text(face = "bold")
  ) 

f17

coef <- all %>%
  filter(TRTA == "Xanomeline High Dose") %>%
  lm(data = ., AVAL ~ BASE) %>%
  coefficients()

coef

all %>%
  filter(TRTA == "Xanomeline High Dose") %>%
  ggplot(., aes(x = BASE, y = AVAL)) +
  geom_point() +
  stat_smooth(method = "lm", se = FALSE, size = 5) +
  geom_abline(intercept = coef[[1]], slope = coef[[2]], color = "red")


# altervatively, plot with patchwork::plot_annotation

# setup plot text
yaxis_text_alt <- "Maximum Post-baseline Systolic Blood Pressure (mmHg)"
xaxis_text_alt <- "Baseline Systolic Blood Pressure (mmHg)"
header_alt <- "Figure 17. Baseline vs. Maximum Systolic Blood Pressure by Treatment Arm, Safety Population¹, Pooled Analysis"
footer1_alt <- "Source: [include Applicant source, datasets and/or software tools used]."
footer2_alt <- "¹Gray dotted line = no increase; blue line = high-dose treatment linear regression; grey dashed line = low-dose linear regression."

f17_alt <- ggplot(all, aes(x = BASE, y = AVAL, linetype = TRTA)) +
  geom_point(colour = "black", shape = 21, size = 4, alpha = 0.6, aes(fill = TRTA)) +
  stat_smooth(method = "lm", se = FALSE, aes(color = TRTA)) +
  scale_color_manual(values = c("skyblue", "grey")) +
  scale_fill_manual(values = c("skyblue", "white")) +
  geom_abline(intercept = 0, slope = 1, size = 1, lty = "dotted", color = "#5A5A5A") +
  scale_x_continuous(breaks = seq(90, 210, 30), limits = c(90, 210)) +
  scale_y_continuous(breaks = seq(90, 210, 30), limits = c(90, 210)) +
  labs(
    y = yaxis_text_alt,
    x = xaxis_text_alt
  ) +
  theme_bw() +
  theme(
    plot.background = element_rect(fill = NA, color = "skyblue3", size = 2, linetype = "solid"),
    legend.position = "bottom",
    legend.title = element_blank(),
    axis.text = element_text(face = "bold")
  ) 

# Use the patchwork package to specify title, caption (e.g. footers) and footers position (i.e. left align) to get a closer aesthetic feel.
library(patchwork)
f17_alt + patchwork::plot_annotation(title = header_alt,
                                     caption = paste(footer1_alt, footer2_alt, sep = "\n"),
                                     theme = theme(plot.caption = element_text(hjust = 0)))


```
# ADAE data frp, PhUSE Test Data Factory repository

```{r r tornade}

# Packages
library(ggplot2)
library(dplyr)
library(ggh4x)

# Data
adae <- haven::read_xpt(
  paste0("https://github.com/phuse-org/TestDataFactory/",
         "raw/main/Updated/TDF_ADaM/adae.xpt"))

# prepare data
start <- adae %>%
  select(TRTA, USUBJID, AESEV, AEBODSYS) %>%
  filter(TRTA != "Placebo")

event_counts <- start %>%
  distinct(USUBJID, AEBODSYS, AESEV, .keep_all = TRUE) %>%
  count(TRTA, AEBODSYS, AESEV)

ggplot(event_counts, aes(fill = AESEV, y = n, x = reorder(AEBODSYS, n))) +
  geom_bar(position="stack", stat="identity") +
  coord_flip() +
  facet_wrap(~ TRTA, scales = "free_x") +
  #facet_wrap(~ TRTA, scales = "free_y") +
  facetted_pos_scales(y = list(
    scale_y_reverse(),
    scale_y_continuous()),
    
  )

## do some customiztion

ggplot(event_counts, aes(fill = AESEV, y = n, x = reorder(AEBODSYS, n))) +
  geom_bar(position="stack", stat="identity", color = "black") +
  coord_flip() +
  scale_fill_brewer(palette = "Blues") +
  facet_wrap(~ TRTA, scales = "free_x") + 
  facetted_pos_scales(y = list(
    scale_y_reverse(breaks = seq(70,0,-10), limits = c(70,0)),
    scale_y_continuous(breaks = seq(0,70,10), limits = c(0,70)))) +
  theme_bw() + 
  theme(legend.position = "bottom",
        legend.title = element_blank(),
        strip.text = element_text(color = "white"),
        strip.background = element_rect(fill = "royalblue"),
        axis.text = element_text(face="bold")) +
  labs(y = "Number of Subjects", x = "System Organ Class")




```

```{r safety}

### load relevant libraries and data 

library(dplyr)
library(haven)
library(patchwork)
library(ggplot2)


adam_path <- "https://github.com/phuse-org/TestDataFactory/raw/main/Updated/TDF_ADaM/"
adsl <- read_xpt(paste0(adam_path, "adsl.xpt"))
adlb <- read_xpt(paste0(adam_path, "adlbc.xpt"))



# prepare lab data set for plotting

data <- adlb %>%
  filter(PARAMCD == "SODIUM" & !is.na(AVISIT)) %>%
  select(USUBJID, TRTA, AVISIT, PARAM, AVAL, BASE, CHG) %>%
  mutate(AVISIT = factor(AVISIT, levels = c("Baseline", "Week 2", "Week 4", "Week 6", "Week 8", "Week 12",
                                            "Week 16", "Week 20", "Week 24", "Week 26", "End of Treatment")))

# create plot

p <- data %>%
  filter(!AVISIT %in% c("Baseline", "End of Treatment") & !is.na(AVISIT)) %>%
  group_by(AVISIT, TRTA) %>%
  mutate(mean_chg = mean(CHG, na.rm = TRUE),
         se_chg = sd(CHG, na.rm = TRUE) / sqrt(length(CHG))) %>%
  ungroup() %>%
ggplot(., aes(x=AVISIT, y=mean_chg, group=TRTA, linetype = TRTA)) +
  geom_errorbar(aes(ymin=mean_chg-se_chg, ymax=mean_chg+se_chg, color=TRTA), width=.1) +  
  geom_line(aes(color=TRTA)) +
  geom_point(aes(color=TRTA)) +
  geom_hline(yintercept=0, linetype='dashed')+
  scale_linetype_manual(values = c("dashed",rep("solid",2))) +
  scale_color_manual(values=c("#93918E","#0B5A8F", "#73BDEE")) +
  labs(x = "", y = "Mean Change from Baseline (95% CI) \n Sodium (mmol/l)") +
  theme_minimal() +
  theme(legend.position="bottom",
        legend.title=element_blank()) +
  coord_fixed() +
  theme(aspect.ratio = 0.5)

p

# prepare data for table 

table_data <- data %>%
  filter(!AVISIT %in% c("Baseline", "End of Treatment") & !is.na(AVISIT)) %>%
  group_by(AVISIT, TRTA) %>%
  mutate(mean_chg = mean(CHG, na.rm = TRUE),
         mean_aval  = mean(AVAL, na.rm  = TRUE),
         n_patients = n_distinct(USUBJID),
          n_patients = n_distinct(USUBJID), 
         TRTA = case_when(
           TRTA == "Xanomeline High Dose" ~"High Dose",
           TRTA == "Xanomeline Low Dose" ~ "Low Dose",
           TRUE ~ TRTA)) %>%
  ungroup() %>%
  select(TRTA, AVISIT, mean_chg, mean_aval, n_patients) %>%
  distinct()

data_table1 <- table_data %>%
  select(-n_patients) %>%
  mutate(value = paste0(round(mean_chg, 2), "/", round(mean_aval, 2)))

data_table2 <- table_data %>%
  select(-mean_chg, -mean_aval)

# plot table 1

p_table1 <- ggplot(data = data_table1, aes(x = AVISIT, y = TRTA)) +
  scale_shape_manual(values = 1:length(data_table1$TRTA))+
  ggpubr::geom_exec(geom_text, data = data_table1, label = data_table1$value, color = "TRTA", size=3) +
  theme_classic() +
  labs(title = "Mean Change from Baseline / Mean Value", x = "", y = "" ) +
  scale_color_manual(values=c("#93918E","#0B5A8F", "#73BDEE")) +
  theme(axis.text.y = element_text(colour = c("#93918E","#0B5A8F", "#73BDEE")),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.ticks.y=element_blank()) +
  theme(legend.position="none",
        panel.border = element_rect(colour = "black", fill=NA, size=1),
        plot.title = element_text(size=10, face = "bold"))


# just for the sake of displaying it

p_table1 +
    coord_fixed(ratio=.5)

## another way to plot

# plot table 2

p_table2 <- ggplot(data = data_table2, aes(x = AVISIT, y = TRTA)) +
  scale_shape_manual(values = 1:length(data_table2$TRTA))+
  ggpubr::geom_exec(geom_text, data = data_table2, label = data_table2$n_patients, color = "TRTA", size=3) +
  theme_classic() +
  labs(title = "Number of Patients with Data", x = "", y = "" ) +
  scale_color_manual(values=c("#93918E","#0B5A8F", "#73BDEE")) +
  theme(axis.text.y = element_text(colour = c("#93918E","#0B5A8F", "#73BDEE")),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.ticks.y=element_blank()) +
  theme(legend.position="none",
        panel.border = element_rect(colour = "black", fill=NA, size=1),
        plot.title = element_text(size=10, face = "bold"))

# just for the sake of displaying it

p_table2 +
    coord_fixed(ratio=.5)


# combine plots

final_plot <- p + p_table1 + p_table2

# display plot w/ layout
final_plot + plot_layout(byrow = FALSE,
                        heights = c(5,1,1))



```







